DisReserv (v1.0): Analytical displacement solution due to reservoir compaction with arbitrary geometry and under arbitrary pressure changes

### 2020-09-20 manuscript/manuscript-reservoir.tex
small change

By following \cite{Sharma56} and \cite{Tempone10}, 
the stress vector 
$\mbox{\boldmath$\sigma$}_{ij} \equiv \mbox{\boldmath$\sigma$}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ 
at the point $(x_{i}, y_{i}, z_{i})$ due to the $j$th single nucleus of strain buried 
in the half space is given by
\begin{equation}
\mbox{\boldmath$\sigma$}_{ij} = \mbox{\boldmath$\sigma_{1}$}_{ij} + 
\mbox{\boldmath$\sigma_{2}$}_{ij} \: ,
\label{eq:stress_nucleus}
\end{equation}
where 
$\mbox{\boldmath$\sigma_{1}$}_{ij} \equiv \mbox{\boldmath$\sigma_{1}$}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ is the stress vector 
at the point $(x_{i}, y_{i}, z_{i})$ due to the $j$th single nucleus in the infinite space and $\mbox{\boldmath$\sigma_{2}$}_{ij} \equiv \mbox{\boldmath$\sigma_{2}$}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ is the stress vector 
at the point $(x_{i}, y_{i}, z_{i})$ that gives the correction of the stress due to the 
$j$th image nucleus considering a semi-space. These two vectors are given by
\begin{equation}
\mbox{\boldmath$\sigma_{1}$}_{ij} =
A  \: 
\begin{bmatrix} 
\frac{\partial^{2} }{\partial x \partial z} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial^{2} }{\partial y \partial z} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial^{2} }{\partial z^{2}} {\frac{1}{{R_1}_{ij}}} 
\end{bmatrix}
\: \Delta p_{j} \: \:dv_j^{\prime} \: ,
\label{eq:nucleus-stress1-vector}
\end{equation}
and 
\begin{equation}
\mbox{\boldmath$\sigma_{2}$}_{ij} = 
A  \: 
\left\{
\begin{bmatrix} 
\frac{\partial^{2}  }{\partial x \partial z} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{2} }{\partial y \partial z} {\frac{1}{{R_2}_{ij}}} \\
- \frac{\partial^{2} }{\partial z^{2}} {\frac{1}{{R_2}_{ij}}}   
\end{bmatrix}
+ 2 \: z_{i}
\begin{bmatrix} 
\frac{\partial^{3}  }{\partial x \partial z^{2}} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{3} }{\partial y \partial z^{2}} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{3} }{\partial z^{3}} {\frac{1}{{R_2}_{ij}}} 
\end{bmatrix}
\right\}
\: \Delta p_{j} \: \:dv_j^{\prime} \: .
\label{eq:nucleus-stress2-vector}
\end{equation}
According to \cite{Sharma56} and \cite{Tempone10}, the Beltrami’s equations \citep{Beltrami} and the equilibrium equations must be satisfied to obtain the contribution of the stress field in the half space. 
Additionally, the following boundary conditions at the free surface ($z_i = 0$) 
must be satisfied
\begin{equation}
\mbox{\boldmath$\sigma$}_{ij} = \mbox{\boldmath$\sigma_{1}$}_{ij} + 
\mbox{\boldmath$\sigma_{2}$}_{ij} = \mbox{ \boldmath$0$},
\label{eq:null_stress}
\end{equation}
where $\mbox{ \boldmath$0$}$ is the null vector.
This condition can be easily verified by adding the vectors 
$\mbox{\boldmath$\sigma_{1}$}_{ij}$ (equation \ref{eq:nucleus-stress1-vector}) and 
$\mbox{\boldmath$\sigma_{2}$}_{ij}$ (equation \ref{eq:nucleus-stress2-vector}) 
computed at any point on the free surface $(x_i, y_i, z_i = 0)$.

@@

In equations \ref{dx1}$-$\ref{dz1} and \ref{dx2}$-$\ref{dz1}, we adopted the modifications proposed by \cite{Fukushima2020}.
To overcome the zero division in evaluating the arguments of the arctangent function, \cite{Fukushima2020} replaced  $\tan^{-1} \big( \frac{S}{T} \big)$ by 
\begin{equation}
arctan2(S,T) = 
  \left\{ \begin{aligned}
      atan (S/T) & \:\: \mbox{if} \: \:\:  T \neq 0 \\
      \pi /2 & \:\: \mbox{if} \: \:\:  T = 0  \: \: \mbox{and} \: \:S > 0 \\
      -\pi /2 & \:\: \mbox{if} \: \:\:  T = 0  \: \: \mbox{and} \: \:S < 0 \\
         0 & \:\: \mbox{if} \: \:\:  T = 0  \: \: \mbox{and} \: \:S = 0 \\
  \end{aligned} \right.
\label{eq:arctan2}  
\end{equation}
If the argument of the logarithm is less than $10^{-10}$, the logarithm is replaced by zero; otherwise the logarithm is  calculated regularly

@@

Figure \ref{fig:Null_stress} shows the  null stress through the free surface at the plane 
$z=0$ m (equations \ref{eq:null_stress} and \ref{eq:null_stress_i}) due to reservoir under uniform depletion.

@@

We verify that the zero stress condition is satisfied at the free surface (equations \ref{eq:null_stress} and \ref{eq:null_stress_i}) due to reservoir under a non uniform depletion as shown in Figure \ref{fig:Null_stress_non_uniform_depletion}.

@@

In this numerical application, the reservoir model is a simplification of a realistic reservoir located in a production oil field in offshore Brazil.
The entire reservoir model comprises dimensions of 14 km in the north-axis, 13 km in the
east-axis, and 0.6 km in the down-axis. 
The depths to the top and bottom of the reservoir model are 2,712 m and 3,312 m, respectively. 
The components of the displacements are calculated at 0 m deep, 
on a regular grid of 100 $\times$ 80  observation points along the north- and east-directions, respectively. 
We  discretized the reservoir  along the $x-$, $y-$ and $z-$ directions into an $14 \times 13 \times 2$ grid of prisms.
The Young’s modulus is  3300 (in MPa), the Poisson's coefficient is 0.25, and
the uniaxial compaction coefficient $C_{m}$  (equation \ref{eq:Cm}) is $2.2525 \: 10^{-4}$
$\textrm{ MPa}^{-1}$.
In the 3D perspective view of the pore pressure distribution of the simulated reservoir shown in Figure \ref{fig:pressure_complex_reservoir}, we can see that the pressures vary from $0$ to $-0.72$ MPa.
Figure \ref{fig:displacement_complex_reservoir} shows cross-sections at $x  = 8$ km of the horizontal and vertical displacements, in 2D contour plots, calculated in the whole reservoir by using our methodology.
Figure \ref{fig:Null_stress_complex_reservoir} shows the  null stress through the free surface (equations \ref{eq:null_stress} and \ref{eq:null_stress_i}) due to reservoir with arbitrary geometry and under arbitrary pressure disribution.

@@

\begin{figure}[h]
\includegraphics[width=8.3cm]{Fig/Figure_Nucleus_Strain.png}
\caption{Schematic representation of the geometry of the nucleus of strain in a semi-infinite medium. After \cite{Munoz&Roehl17}. 
The adopted Cartesian coordinate system considered the $x-$axis pointing to north, the $y-$axis pointing to east and the $z-$axis pointing downward.} 
\label{fig:nucleus_strain}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=8.3cm]{Fig/Figure_Cylinder.png}
\caption{Disk-shaped reservoir under uniform depletion with a radius of 500 m}
\label{fig:cylinder}
\end{figure}

\begin{figure}[h]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement.png}
\caption{Reservoir under uniform depletion: (a) Horizontal displacement (equation \ref{eq:horizontal_displacement}) and (b) vertical displacement (equation \ref{eq:u_til_z}) by our methodology that uses the closed expressions of the volume integrations given by \cite{Nagyetal2000} and \cite{Nagyetal2002} (equations \ref{dx1}-\ref{dz2})}
\label{fig:displacement}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement_z_levels.png}
\caption{Reservoir under uniform depletion: (a) Horizontal x-component displacement and (b) vertical displacement by our methodology that uses the closed expressions of the volume integrations given by \cite{Nagyetal2000} and \cite{Nagyetal2002} (equations \ref{dx1}-\ref{dz2}).
These displacements are calculated along the x-axis, at $y = 0$ m and $z$ located at the depths of:  seafloor ($z = 0$ m), reservoir top ($z = 750$ m), reservoir center ($z = 800$ m) and reservoir bottom ($z = 850$ m).}
\label{fig:displacement_z_levels}
\end{figure}

\begin{figure}[h]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement_z_levels_Geertsma.png}
\caption{Reservoir under uniform depletion: Vertical displacement using Geertsma’s method \citep{Geertsma73}  considering an elastic homogeneous cylindrical reservoir under uniform depletion \citep{Fjaer08}.
The displacement is calculated along the x-axis, at $y = 0$ m and $z$ located at the depths of:  seafloor ($z = 0$ m), reservoir top ($z = 750$ m), and reservoir bottom 
($z = 850$ m).}
\label{fig:displacement_z_levels_Geertsma}
\end{figure}

\begin{figure*}[h]
\includegraphics[width=12cm]{Fig/Figure_Null_stress.png}
\caption{Reservoir under uniform depletion: (a) $x-$, (b) $y-$, and (c) $z-$ components of the stress at the free surface.}
\label{fig:Null_stress}
\end{figure*}

@@

\begin{figure}[h]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement_non_uniform_depletion.png}
\caption{Reservoir under non uniform depletion: (a) Horizontal displacement (equation \ref{eq:horizontal_displacement}) and (b) vertical displacement (equation \ref{eq:u_til_z}) by our methodology that uses the closed expressions of the volume integrations given by \cite{Nagyetal2000} and \cite{Nagyetal2002} (equations \ref{dx1}-\ref{dz2})}
\label{fig:displacement_non_uniform_depletion}
\end{figure}

\begin{figure}[h]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement_z_levels_non_uniform_depletion.png}
\caption{Reservoir under non uniform depletion: (a) Horizontal x-component displacement and (b) vertical displacement by our methodology that uses the closed expressions of the volume integrations given by \cite{Nagyetal2000} and \cite{Nagyetal2002} (equations \ref{dx1}-\ref{dz2}).
These displacements are calculated along the x-axis, at $y = 0$ m and $z$ located at the depths of:  seafloor ($z = 0$ m), reservoir top ($z = 750$ m), reservoir center ($z = 800$ m) and reservoir bottom ($z = 850$ m).}
\label{fig:displacement_z_levels_non_uniform_depletion}
\end{figure}

\begin{figure*}[h]
\includegraphics[width=12cm]{Fig/Figure_Null_stress_non_uniform_depletion.png}
\caption{Reservoir under non uniform depletion: (a) $x-$, (b) $y-$, and (c) $z-$ components of the stress at the free surface.}
\label{fig:Null_stress_non_uniform_depletion}
\end{figure*}

@@

\begin{figure}[h]
\includegraphics[width=8.3cm]{Fig/Figure_Pressure_complex_reservoir.png}
\caption{Reservoir with arbitrary geometry and under arbitrary pressure changes: 3D perspective view of the pore pressure distribution of a realistic reservoir 
located in a production oil field in offshore Brazil.}
\label{fig:pressure_complex_reservoir}
\end{figure}

\begin{figure}[h]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement_complex_reservoir.png}
\caption{Reservoir with arbitrary geometry and under arbitrary pressure changes: (a) Horizontal displacement (equation \ref{eq:horizontal_displacement}) and (b) vertical displacement (equation \ref{eq:u_til_z}) by our methodology that uses the closed expressions of the volume integrations given by \cite{Nagyetal2000} and \cite{Nagyetal2002} (equations \ref{dx1}-\ref{dz2})}
\label{fig:displacement_complex_reservoir}
\end{figure}

\begin{figure*}[h]
\includegraphics[width=12cm]{Fig/Figure_Null_stress_complex_reservoir.png}
\caption{Reservoir with arbitrary geometry and under arbitrary pressure changes: (a) $x-$, (b) $y-$, and (c) $z-$ components of the stress at the free surface.}
\label{fig:Null_stress_complex_reservoir}
\end{figure*}

>>>>

By following \cite{Sharma56} and \cite{Tempone10}, 
the stress vector 
$\mbox{\boldmath$\sigma$}_{ij} \equiv \mbox{\boldmath$\sigma$}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ 
at the point $(x_{i}, y_{i}, z_{i})$ due to the $j$th single nucleus of strain buried 
in the half space is given by
\begin{equation}
\mbox{\boldmath$\sigma$}_{ij} = \mbox{\boldmath$\sigma_{1}$}_{ij} + 
\mbox{\boldmath$\sigma_{2}$}_{ij} \: ,
\label{eq:stress_nucleus}
\end{equation}
where 
$\mbox{\boldmath$\sigma_{1}$}_{ij} \equiv \mbox{\boldmath$\sigma_{1}$}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ is the stress vector 
at the point $(x_{i}, y_{i}, z_{i})$ due to the $j$th single nucleus in the infinite space and $\mbox{\boldmath$\sigma_{2}$}_{ij} \equiv \mbox{\boldmath$\sigma_{2}$}(x_{i}, y_{i}, z_{i}, x^{\prime}_{j}, y^{\prime}_{j}, z^{\prime}_{j})$ is the stress vector 
at the point $(x_{i}, y_{i}, z_{i})$ that gives the correction of the stress due to the 
$j$th image nucleus considering a semi-space. These two vectors are given by
\begin{equation}
\mbox{\boldmath$\sigma_{1}$}_{ij} =
A  \: 
\begin{bmatrix} 
\frac{\partial^{2} }{\partial x \partial z} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial^{2} }{\partial y \partial z} {\frac{1}{{R_1}_{ij}}}  \\
\frac{\partial^{2} }{\partial z^{2}} {\frac{1}{{R_1}_{ij}}} 
\end{bmatrix}
\: \Delta p_{j} \: \:dv_j^{\prime} \: ,
\label{eq:nucleus-stress1-vector}
\end{equation}
and 
\begin{equation}
\mbox{\boldmath$\sigma_{2}$}_{ij} = 
A  \: 
\left\{
\begin{bmatrix} 
\frac{\partial^{2}  }{\partial x \partial z} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{2} }{\partial y \partial z} {\frac{1}{{R_2}_{ij}}} \\
- \frac{\partial^{2} }{\partial z^{2}} {\frac{1}{{R_2}_{ij}}}   
\end{bmatrix}
+ 2 \: z_{i}
\begin{bmatrix} 
\frac{\partial^{3}  }{\partial x \partial z^{2}} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{3} }{\partial y \partial z^{2}} {\frac{1}{{R_2}_{ij}}} \\
\frac{\partial^{3} }{\partial z^{3}} {\frac{1}{{R_2}_{ij}}} 
\end{bmatrix}
\right\}
\: \Delta p_{j} \: \:dv_j^{\prime} \: .
\label{eq:nucleus-stress2-vector}
\end{equation}
According to \cite{Sharma56} and \cite{Tempone10}, the Beltrami’s equations \citep{Beltrami} and the equilibrium equations must be satisfied to obtain the contribution of the stress field in the half space. 
Additionally, the boundary condition 
$\mbox{\boldmath$\sigma$}_{ij} = \mbox{ \boldmath$0$}$ (equation \ref{eq:stress_nucleus})
must be satisfied at the free surface ($z_i = 0$).
%\begin{equation}
%\mbox{\boldmath$\sigma$}_{ij} = \mbox{\boldmath$\sigma_{1}$}_{ij} + 
%\mbox{\boldmath$\sigma_{2}$}_{ij} = \mbox{ \boldmath$0$},
%\label{eq:null_stress}
%\end{equation}
%where $\mbox{ \boldmath$0$}$ is the null vector.
This condition can be easily verified by adding the vectors 
$\mbox{\boldmath$\sigma_{1}$}_{ij}$ (equation \ref{eq:nucleus-stress1-vector}) and 
$\mbox{\boldmath$\sigma_{2}$}_{ij}$ (equation \ref{eq:nucleus-stress2-vector}) 
computed at any point on the free surface $(x_i, y_i, z_i = 0)$.

@@

In equations \ref{dx1}--\ref{dz1} and \ref{dx2}--\ref{dz1}, we adopted the modifications proposed by \cite{Fukushima2020}.
To overcome the zero division in evaluating the arguments of the arctangent function, \cite{Fukushima2020} replaced  $\tan^{-1} \big( \frac{S}{T} \big)$ by 
\begin{equation}
arctan2(S,T) = \begin{cases}
	%\left\{ \begin{aligned}
    atan (S/T) & \:\: \mbox{if} \: \:\:  T \neq 0 \\
    \pi /2 & \:\: \mbox{if} \: \:\:  T = 0  \: \: \mbox{and} \: \:S > 0 \\
    -\pi /2 & \:\: \mbox{if} \: \:\:  T = 0  \: \: \mbox{and} \: \:S < 0 \\
    0 & \:\: \mbox{if} \: \:\:  T = 0  \: \: \mbox{and} \: \:S = 0 \\
    %\end{aligned} \right\}.
    \end{cases}
\label{eq:arctan2}  
\end{equation}
If the argument of the logarithm is less than $10^{-10}$, the logarithm is replaced by zero; otherwise the logarithm is  calculated regularly

@@

Figure \ref{fig:Null_stress} shows the null stress through the free surface at the 
plane $z=0$ m due to reservoir under uniform depletion.

@@

We verify that the zero stress condition is satisfied at the free surface 
%(equations \ref{eq:null_stress} and \ref{eq:null_stress_i}) 
due to reservoir under a non uniform depletion as shown in Figure \ref{fig:Null_stress_non_uniform_depletion}.

@@

In this numerical application, the reservoir model is a simplification of a realistic reservoir located in a production oil field in offshore Brazil.
The entire reservoir model comprises dimensions of 14 km in the north-axis, 13 km in the
east-axis, and 0.6 km in the down-axis. 
The depths to the top and bottom of the reservoir model are 2,712 m and 3,312 m, respectively. 
The components of the displacements are calculated at 0 m deep, 
on a regular grid of 100 $\times$ 80  observation points along the north- and east-directions, respectively. 
We  discretized the reservoir  along the $x-$, $y-$ and $z-$ directions into an $14 \times 13 \times 2$ grid of prisms.
The Young’s modulus is  3300 (in MPa), the Poisson's coefficient is 0.25, and
the uniaxial compaction coefficient $C_{m}$  (equation \ref{eq:Cm}) is $2.2525 \: 10^{-4}$
$\textrm{ MPa}^{-1}$.
In the 3D perspective view of the pore pressure distribution of the simulated reservoir shown in Figure \ref{fig:pressure_complex_reservoir}, we can see that the pressures vary from $0$ to $-0.72$ MPa.
Figure \ref{fig:displacement_complex_reservoir} shows cross-sections at $x  = 8$ km of the horizontal and vertical displacements, in 2D contour plots, calculated in the whole reservoir by using our methodology.
Figure \ref{fig:Null_stress_complex_reservoir} shows the  null stress through the free surface 
%(equations \ref{eq:null_stress} and \ref{eq:null_stress_i}) 
due to reservoir with arbitrary geometry and under arbitrary pressure disribution.

@@

\begin{figure}[ht]
\includegraphics[width=8.3cm]{Fig/Figure_Nucleus_Strain.png}
\caption{Schematic representation of the geometry of the nucleus of strain in a semi-infinite medium. After \cite{Munoz&Roehl17}. 
The adopted Cartesian coordinate system considered the $x-$axis pointing to north, the $y-$axis pointing to east and the $z-$axis pointing downward.} 
\label{fig:nucleus_strain}
\end{figure}

@@

\begin{figure}[ht]
\includegraphics[width=8.3cm]{Fig/Figure_Cylinder.png}
\caption{Disk-shaped reservoir under uniform depletion with a radius of 500 m}
\label{fig:cylinder}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement.png}
\caption{Reservoir under uniform depletion: (a) Horizontal displacement (equation \ref{eq:horizontal_displacement}) and (b) vertical displacement (equation \ref{eq:u_til_z}) by our methodology that uses the closed expressions of the volume integrations given by \cite{Nagyetal2000} and \cite{Nagyetal2002} (equations \ref{dx1}-\ref{dz2})}
\label{fig:displacement}
\end{figure}

@@

\begin{figure}[ht]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement_z_levels.png}
\caption{Reservoir under uniform depletion: (a) Horizontal x-component displacement and (b) vertical displacement by our methodology that uses the closed expressions of the volume integrations given by \cite{Nagyetal2000} and \cite{Nagyetal2002} (equations \ref{dx1}-\ref{dz2}).
These displacements are calculated along the x-axis, at $y = 0$ m and $z$ located at the depths of:  seafloor ($z = 0$ m), reservoir top ($z = 750$ m), reservoir center ($z = 800$ m) and reservoir bottom ($z = 850$ m).}
\label{fig:displacement_z_levels}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement_z_levels_Geertsma.png}
\caption{Reservoir under uniform depletion: Vertical displacement using Geertsma’s method \citep{Geertsma73}  considering an elastic homogeneous cylindrical reservoir under uniform depletion \citep{Fjaer08}.
The displacement is calculated along the x-axis, at $y = 0$ m and $z$ located at the depths of:  seafloor ($z = 0$ m), reservoir top ($z = 750$ m), and reservoir bottom 
($z = 850$ m).}
\label{fig:displacement_z_levels_Geertsma}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=12cm]{Fig/Figure_Null_stress.png}
\caption{Reservoir under uniform depletion: (a) $x-$, (b) $y-$, and (c) $z-$ components of the stress at the free surface.}
\label{fig:Null_stress}
\end{figure}

@@

\begin{figure}[ht]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement_non_uniform_depletion.png}
\caption{Reservoir under non uniform depletion: (a) Horizontal displacement (equation \ref{eq:horizontal_displacement}) and (b) vertical displacement (equation \ref{eq:u_til_z}) by our methodology that uses the closed expressions of the volume integrations given by \cite{Nagyetal2000} and \cite{Nagyetal2002} (equations \ref{dx1}-\ref{dz2})}
\label{fig:displacement_non_uniform_depletion}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement_z_levels_non_uniform_depletion.png}
\caption{Reservoir under non uniform depletion: (a) Horizontal x-component displacement and (b) vertical displacement by our methodology that uses the closed expressions of the volume integrations given by \cite{Nagyetal2000} and \cite{Nagyetal2002} (equations \ref{dx1}-\ref{dz2}).
These displacements are calculated along the x-axis, at $y = 0$ m and $z$ located at the depths of:  seafloor ($z = 0$ m), reservoir top ($z = 750$ m), reservoir center ($z = 800$ m) and reservoir bottom ($z = 850$ m).}
\label{fig:displacement_z_levels_non_uniform_depletion}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=12cm]{Fig/Figure_Null_stress_non_uniform_depletion.png}
\caption{Reservoir under non uniform depletion: (a) $x-$, (b) $y-$, and (c) $z-$ components of the stress at the free surface.}
\label{fig:Null_stress_non_uniform_depletion}
\end{figure}

@@

\begin{figure}[ht]
\includegraphics[width=8.3cm]{Fig/Figure_Pressure_complex_reservoir.png}
\caption{Reservoir with arbitrary geometry and under arbitrary pressure changes: 3D perspective view of the pore pressure distribution of a realistic reservoir 
located in a production oil field in offshore Brazil.}
\label{fig:pressure_complex_reservoir}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=8.3cm]{Fig/Figure_Displacement_complex_reservoir.png}
\caption{Reservoir with arbitrary geometry and under arbitrary pressure changes: (a) Horizontal displacement (equation \ref{eq:horizontal_displacement}) and (b) vertical displacement (equation \ref{eq:u_til_z}) by our methodology that uses the closed expressions of the volume integrations given by \cite{Nagyetal2000} and \cite{Nagyetal2002} (equations \ref{dx1}-\ref{dz2})}
\label{fig:displacement_complex_reservoir}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=12cm]{Fig/Figure_Null_stress_complex_reservoir.png}
\caption{Reservoir with arbitrary geometry and under arbitrary pressure changes: (a) $x-$, (b) $y-$, and (c) $z-$ components of the stress at the free surface.}
\label{fig:Null_stress_complex_reservoir}
\end{figure}
